Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-28 cis-perf-new-ll.tex

Otherwise, the sets are defined such that $\nu_j$ and $\nu_{j'}$ are in the same class if and only if $i(j) = i(j')$ and the episodes beginning and end times are censored to the same interval.
If $i \in \set{D}$ then there is exactly one $k$ such that $\set{E}_k$ is identical to $i$'s admissible region $\alpha_i$.
% In this case, I write $\alpha_{i(k)}$ to denote this admissible region.

@@

Let $\xi_k$ denote $\prob(\nu_j \in \set{E}_k)$.
Assume, for tractability, that the events $\nu_j \in \set{E}_k$ and $\nu_{j'} \in \set{E}_{k'}$ are independent for $j \neq j'$.
Denote by $m_k$ then number of episodes in $\set{E}_k$.
$\vec{m} = [m_1, \dots m_{N_E}, m_u]^T$ are the observations.
Then:
\begin{align}
  \vec{m}
  \mid \ntot, \vec{\theta}
  \dist
  \MN\left(
    \ntot,
    \frac{1}{\Ncis}
    \begin{bmatrix}
        \xi_1 \\ \xi_2 \\ \vdots \\ \xi_{N_E} \\ \xi_u
    \end{bmatrix}
  \right).
  \label{perf-test:eq:alt-likelihood}
\end{align}

>>>>

Otherwise, the sets are defined such that $\nu_j$ and $\nu_{j'}$ are in the same class if and only if $i(j) = i(j')$ and the episodes beginning and end times are censored to be in the same pair of intervals.
If $i \in \set{D}$ then there is exactly one $k$ such that $\set{E}_k$ is identical to $i$'s admissible region $\alpha_i$.
% In this case, I write $\alpha_{i(k)}$ to denote this admissible region.

@@

Let $\xi_k$ denote $\prob(\nu_j \in \set{E}_k)$.
Assume, for tractability, that the events $\nu_j \in \set{E}_k$ and $\nu_{j'} \in \set{E}_{k'}$ are independent for $j \neq j'$.
Denote by $m_k$ the number of episodes in $\set{E}_k$.
$\vec{m} = [m_1, \dots m_{N_E}, m_u]^T$ are the observations.
Then:
\begin{align}
  \vec{m}
  \mid \ntot, \vec{\theta}
  \dist
  \MN\left(
    \ntot,
    \frac{1}{\Ncis}
    \begin{bmatrix}
        \xi_1 \\ \xi_2 \\ \vdots \\ \xi_{N_E} \\ \xi_u
    \end{bmatrix}
  \right).
  \label{perf-test:eq:alt-likelihood}
\end{align}

### 2024-03-28 incidence-prevalence.tex
ch 7 proofing

\section{Infection process} \label{inc-prev:sec:infection-process}
The infection process is the mathematical description of how infections occur.
It is a \emph{point} or \emph{counting} process, where each infection is an event.

A point process is a random function of continuous time $K(t)$.
$K(t)$ must take non-negative integer values and be monotonically increasing in $t$.
The interpretation of $K(t)$ is the number of events, here infections, which have occurred in the interval $(-\infty, t]$~\autocite[244]{yanDistribution}.
In the context of a disease, the point process can be discretised into the incidence on day $t$ as $Z_t = K(t) - K(t-1)$, where $t$ is an integer.
I will explain two approaches to modelling the infection process.

@@

A commonly used point process for infectious disease epidemiology is the time inhomogeneous Poisson process~\autocite[e.g., in the context of HIV,][]{brookmeyerMethod,paganoHIV,rosenbergBackcalculation,brookmeyerBackcalculation}.
% A time-homogenous Poisson process is the special case where $\nu(r)$ is constant.
Extensions to this model, which I do not consider here, have incorporated covariates, such as geography or age~\autocite[e.g.][]{diggleModeling}.

@@

Considering multiple values of $t$ makes \cref{inc-prev:eq:EPt} into a system of simultaneous equations.
Therefore, the system can be written in matrix form:
\begin{align}
  \E(\vec{P} \mid \vec{Z}) = \matr{S} \vec{Z}. \label{inc-prev:eq:EPt-matrix} \\
\intertext{$\matr{S}$ is a lower-triangular matrix formed from $S$:}
    (\matr{S})_{i,j} &= \begin{cases}
        S(i - j + 1) & i - \dmax - 1 \leq j \leq i \\
        0 & \text{otherwise.}
    \end{cases} \label{inc-prev:eq:Smatrix}
\end{align}
In \cref{E-backcalc}, I will need to solve for this system for $\vec{Z}$ given a known $\hat{P} = \E(\vec{P} \mid {Z})$.
Assuming $\vec{P}$ and $\vec{Z}$ have the same dimension then $\matr{S}$ is lower-triangular with strictly positive diagonal elements; therefore, $\matr{S}$ is invertible, and the system has a unique solution~\autocite[424]{bernsteinFacts}.
The lower-triangular structure of $\matr{S}$ makes allows solving the system through forward-substitution, an efficient method that avoids explicit matrix inversion~\autocite{cormenMatrix}.
Forward-substitution sequentially solves the $t$th element of $\vec{Z}$ ($t = 1, \dots, T$) by calculating:
\begin{align}
Z_t
&= \frac{\hat{P}_t - \sum_{i=1}^{t-1} S_{ti} Z_i}{S_{tt}} \\
&= \hat{P}_t - \sum_{i=1}^{t-1} S(t - i + 1) Z_i.
\label{inc-prev:eq:forward-substitute}
\end{align}

>>>>

\section{Infection process} \label{inc-prev:sec:infection-process}
The infection process is the mathematical description of how infections occur.
It is a \emph{counting} process, where each infection is an event.

A counting process is a random function of continuous time $K(t)$.
$K(t)$ must take non-negative integer values and be monotonically increasing in $t$.
The interpretation of $K(t)$ is the number of events, here infections, which have occurred in the interval $(-\infty, t]$~\autocite[244]{yanDistribution}.
In the context of a disease, the counting process can be discretised into the incidence on day $t$ as $Z_t = K(t) - K(t-1)$, where $t$ is an integer.
I will explain two approaches to modelling the infection process.

@@

A commonly used counting process for infectious disease epidemiology is the time inhomogeneous Poisson process~\autocite[e.g., in the context of HIV,][]{brookmeyerMethod,paganoHIV,rosenbergBackcalculation,brookmeyerBackcalculation}.
% A time-homogenous Poisson process is the special case where $\nu(r)$ is constant.
Extensions to this model, which I do not consider here, have incorporated covariates, such as geography or age~\autocite[e.g.][]{diggleModeling}.

@@

Considering multiple values of $t$ makes \cref{inc-prev:eq:EPt} into a system of simultaneous equations.
Therefore, the system can be written in matrix form:
\begin{align}
  \E(\vec{P} \mid \vec{Z}) = \matr{S} \vec{Z}. \label{inc-prev:eq:EPt-matrix} \\
\intertext{$\matr{S}$ is a lower-triangular matrix formed from $S$:}
    (\matr{S})_{i,j} &= \begin{cases}
        S(i - j + 1) & i - \dmax - 1 \leq j \leq i \\
        0 & \text{otherwise.}
    \end{cases} \label{inc-prev:eq:Smatrix}
\end{align}
In \cref{E-backcalc}, I will need to solve for this system for $\vec{Z}$ given a known $\hat{P} = \E(\vec{P} \mid {Z})$.
Assuming $\vec{P}$ and $\vec{Z}$ have the same dimension then $\matr{S}$ is lower-triangular with strictly positive diagonal elements; therefore, $\matr{S}$ is invertible, and the system has a unique solution~\autocite[424]{bernsteinFacts}.
The lower-triangular structure of $\matr{S}$ makes allows solving the system through forward-substitution, an efficient method that avoids explicit matrix inversion~\autocite{cormenMatrix}.
Forward-substitution sequentially solves the $t$th element of $\vec{Z}$ ($t = 1, \dots, T$) by calculating:
\begin{align}
Z_t
&= \frac{\hat{P}_t - \sum_{i=1}^{t-1} S_{ti} Z_i}{S_{tt}} \\
&= \hat{P}_t - \sum_{i=1}^{t-1} S(t - i + 1) Z_i.
\label{inc-prev:eq:forward-substitute}
\end{align}
where $S(t)$ for $t > \dmax$ is defined to be 0.
